### 2018-03-07 Draft/Chapter2/background.tex

\chapter{Background}

\section{Cosmic fields}
Table from the from talk
DTFE, Hessians etc - structured, unstructured etc

\section{Zel'dovich Approximation, scales}
Stuff from Cosmology notes

\section{Dynamics: Lagrangian sub-manifold}
Multi-stream flows, Caustics, Flip-flips