### 2018-03-07 Draft/thesis-ku.tex

%% LyX 2.0.3 created this file.  For more info, see http://www.lyx.org/.
%% Do not edit unless you really know what you are doing.
\documentclass[12pt,english]{kuthesis}

\usepackage{amssymb}
\usepackage{amsmath}   % Added by Nesar - amsmath creates some issues
\usepackage{amstext}

\usepackage{mathptmx}
\renewcommand{\sfdefault}{lmss}
\renewcommand{\ttdefault}{lmtt}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage{listings}
\usepackage{geometry}
\geometry{verbose,tmargin=1in,bmargin=1in,lmargin=1in,rmargin=1in}
\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}
\usepackage{url}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{esint}   %- Commented by Nesar - check conflict with amsmath
\usepackage[authoryear]{natbib}

% Following packages added by Nesar
%\usepackage{amsmath}
\usepackage{epsf}
%\usepackage{graphicx}
\usepackage{graphics}
\usepackage{epsfig}
\usepackage{wrapfig}
\usepackage{listings}
\usepackage{float}
%\usepackage{esint}
\usepackage{natbib}
\usepackage{ifthen}
\usepackage{hyperref}
\usepackage{verbatim}
\usepackage{bm}

% ----------HIGHLIGHTING -------
\usepackage{xcolor}
\usepackage[normalem]{ulem} % use normalem to protect \emph
\newcommand\hl{\bgroup\markoverwith
  {\textcolor{yellow}{\rule[-.5ex]{2pt}{2.5ex}}}\ULon}

\newcommand\rhl{\bgroup\markoverwith
  {\textcolor{orange}{\rule[-.5ex]{2pt}{2.5ex}}}\ULon}

 \newcommand\ghl{\bgroup\markoverwith
  {\textcolor{green}{\rule[-.5ex]{2pt}{2.5ex}}}\ULon} 
%-------------------------------------------------------------------------

%\newcommand{\bmath}[1]{\ensuremath{\boldmath{#1}}\xspace}

\doublespacing

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
\providecommand{\LyX}{L\kern-.1667em\lower.25em\hbox{Y}\kern-.125emX\@}
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% User specified LaTeX commands.

%used to align decimals in tables according to APA style
\usepackage{dcolumn}
\usepackage{booktabs}

% Set the title and author info
\title{Please Read the Abstract}
\author{Nesar Ramachandra}

\dept{Department of People who read Abstracts}
\degreetitle{Doctor of Philosophy}
\papertype{Dissertation} %capitalization is important here
\committee{Sergei Shandarin}{Hume Feldman}{Barbary Anthony-Twarog}{Mikhail Medvedev}{Suzanne Shontz}
%AT Most 5 members allowed, last here is blank on purpose to demonstrate
%flexibility

%% These settings are now in the kuthesis.cls file, but users are free
% to customize. listings has great documentation online
%% When listings are used, break lines
%\lstset{
 %    breaklines=true,  % sets automatic line breaking
 %    breakindent=2em,
 %    breakatwhitespace=true,  % sets if automatic breaks should
 %   breakautoindent=true
%}

%% The following is OPTIONAL. Remove all 3 of the next 3 lines
%% to leave dates blank. If dates are included, then both dates 
%% must be included.
\@printd@testrue
\datedefended{October 02, 2012}
\dateapproved{October 03, 2012}

\@ifundefined{showcaptionsetup}{}{%
 \PassOptionsToPackage{caption=false}{subfig}}
\usepackage{subfig}
\makeatother

\usepackage{babel}
\begin{document}
\begin{romanpages}

\maketitle
\begin{abstract}
Version 2.1, May 20, 2013. This was a minor update that slightly changed
kuthesis.cls. First, the Graduate School no longer exists, so we rephrase
the front matter to refer to Graduate Faculty. Second, the page numbering
after the front matter begins at arabic number 1. This was an oversight
in the previous edition, it was never intended, but users found easy
``work arounds'' and we were not aware of it.

Version 2, October 2, 2012. This is a minor update to allow insertion
of dates in the document preamble.

If you are reading the pdf, this document is produced from ``thesis-ku.tex''
using the \LaTeX{} class kuthesis.cls and other common packages that
are almost surely installed in a working \LaTeX{} system. We suggest
you try to compile this document in your system (as a test), and after
that use it as a template for your project.

\textbf{Look here.} The \LaTeX{} class ``kuthesis.cls'' extends
the base class called ``report''. If you have a report document
already, migration into this format will be easy, since we inherit
everything from ``report'' and then customize a few things.

\textbf{How To Use? }This is a basic \LaTeX{} usage question. You
need a ``\LaTeX{} distribution'' such as TexLive or MikTex. You
need to run the file thesis-ku.tex through the \LaTeX{} processing
system. At the current time, the preferred output format is pdf from
the ``pdflatex'' program, but it is probably safe to try other programs
as well. On most \LaTeX{} systems, your latex processor will look
in several places for ``kuthesis.cls'' while it processes thesis-ku.tex.
Usually, you don't need to install formally that file, since--on most
systems--the \LaTeX{} program will look in the current working directory
for syle files. If so, then this document will ``just work'' for
you. If you want to use kuthesis.cls for several documents, you need
to either copy it into every document directory, or install ``kuthesis.cls''
in a place where your \LaTeX{} system will find it, and then run the
\LaTeX{} re-configurator to make the system notice. In my system,
\LaTeX{} is set up to look in /usr/share/texmf/tex/latex, /usr/local/share/texmf/tex/latex,
as well as in my home directory and the current working directory.
I copy kuthesis.sty into /usr/local/share/texmf/tex/latex, run the
re-configurator program ``texhash,'' and it is ready to go.

\textbf{Does it work with Lyx?} Yes, as a matter of fact, I'm preparing
this document in \LyX{} right now. (\LyX{} is a GUI \LaTeX{} document
editor, available for free at \url{http://www.lyx.org}.) \LyX{} will
open the document, but it will warn that the layout file ``kuthesis.layout''
is unavailable. Unfortunately, the only fix for that is to install
the customized layout file, which is provided with this document in
the stylefiles directory. Copy the kuthesis.layout file into the one
of the folders where \LyX{} looks for layout configurations. I did
not install that in the system-wide \LyX{} layout folder, but rather
in my personal user configuration folder. I suggest you so the same,
at least while you are testing this. On Linux, that is in the user
HOME under .lyx/layouts. (For me, that's ``/home/pauljohn/.lyx/layouts.''
On other OS, here's how you can find out where to put the layout.
Run \LyX{}, in the top menu choose Tools/Preferences. Change something
trivial, close. That will force \LyX{} to create a user folder somewhere
in your personal HOME directories. In MS Windows 7, I just tested
this. I searched for the word ``layouts'' and found my \LyX{} preference
folder is ``C:\textbackslash{}Users\textbackslash{}pauljohn\textbackslash{}AppData\textbackslash{}Roaming\textbackslash{}lyx202\textbackslash{}layouts''.

\textbf{Note.} To customize this document, you have to edit the document
preamble to put in your name, title, degree name, department, and
advisors. If you don't do that, you won't get the right title page.
To be explicit, in this document, the following values are currently
set in the preamble:

\begin{minipage}[t]{1\columnwidth}%
\inputencoding{latin9}\begin{lstlisting}
\title{Please Read the Abstract} 
\author{Abstract Writer} 
\dept{Department of People who read Abstracts} 
\degreetitle{Doctor of Philosophy} 
\papertype{Dissertation} %capitalization is important here 
\committee{MEMBER 1}{MEMBER 2}{MEMBER 3}{MEMBER 4}{}
\end{lstlisting}
\inputencoding{utf8}%
\end{minipage}

\textbf{Note please.} Replace ``Member 1'' with a name, of course.
You can have at most 5 committee members, but you can have fewer.
For the ones you have, put their names inside squiggles; for the ones
you don't want, leave empty squiggles, \{\}, as I've demonstrated
on the 5th member in this example.

\textbf{Note also. }The chapters included here are just for illustration
and testing. I've fiddled the format to look like what KU asks for.

\textbf{Note also. }Note the file ``thesis-ku.tex'' (or thesis-ku.lyx,
as the case may be), does not have much content. It has the title
page, the abstract, and not much else. Everything else is brought
in from separate files. I strongly recommend you keep separate chapters
in separate folders. If you don't do that, you will likely go insane.

\textbf{Note for the version 2012-10-02.} In the first edition of
this setup, I left the dates blank and expected users to write or
type them in as necessary. A user requested a fix to allow the dates
of the defense and final approval to be added in the document preamble
with the other settings. So in the end of the document preamble, insert
the following. The first is some \LaTeX{} code, don't change it.

\begin{minipage}[t]{1\columnwidth}%
\inputencoding{latin9}\begin{lstlisting}
\@printd@testrue
\datedefended{October 2, 2012}
\dateapproved{October 3, 2012}
\end{lstlisting}
\inputencoding{utf8}%
\end{minipage}

\textbf{Note finally.} I want you to use \LaTeX{} to prepare documents.
Doing anything else is ``finger painting.'' But don't ask me for
much help. If you find something wrong with kuthesis.cls, I will be
glad to try to fix it if you let me know. I don't have time/ability
to answer a million email questions about \LaTeX{} or \LyX{}. So please
consult the WWW, there is a ton of stuff up there, that's how I learned.
And if you get in a serious hole and ask me for help, it might cost
you a t-shirt.

Paul Johnson <pauljohn@ku.edu>

Prof. Political Science \& Assoc. Dir., Center for Research Methods
and Data Analysis

2012-10-02

And that's the end of the abstract.

\begin{acknowledgementslong}
%%if you want a "quote" environment for acknowledgements,
%% use acknowledgements instead of acknowledgementslong

I would like to thank mum, dad, Prof. Shandarin, committee members, officemates, friends, exes, uncles, aunts, grandpa, grandma, grandma's dogs, neighbors, neighbors' dogs who made this thesis possible.

\end{acknowledgementslong}

\end{abstract}
\tableofcontents{}

\listoffigures

\listoftables

\end{romanpages}

\include{Chapter1/intro}
\include{Chapter2/background}
\include{Chapter3/Paper2015}
\include{Chapter4/Paper2017a_Nobmath}
\include{Chapter5/Paper2017b_Nobmath}

\global\long\def\bibname{References}

\bibliographystyle{apalike2}
%\bibliography{Biblio/allcites}
\bibliography{Biblio/library2018}

\appendix
\include{Appendix1/Appendix1}
\include{Appendix2/Appendix_paper2017a}
\include{Appendix3/Appendix_paper2017b}

\end{document}