### 2018-05-21 Draft/Chapter3b/nuza.tex

\subsection{CLASSIC\\ \hskip 0.75cm (Manti, Nuza \& Kitaura)}
\label{section:CLASSIC}

The CLASSIC approach is based on performing a prior linearization to the cosmological density field and later a cosmic web classification of the resulting matter distribution. The method is implemented in two steps: first, a linearization is made to better fulfil the mathematical conditions of the original idea of cosmic web classification, which is based on a linear Taylor expansion of the gravitational field \citep[see][]{1974FizSz..24..304Z,hahn2007a}, and then, cosmological structures are divided into voids, sheets, filaments and knots. The linearization is done using higher order Lagrangian perturbation theory as proposed by \cite{2012MNRAS.425.2443K}. In this framework, a given density field can be expressed as the sum of a linear and a non-linear component which are tightly coupled to each other by the tidal field tensor. The cosmic web classification is performed on a grid cell in a similar way as suggested by \cite{hahn2007a}, i.e. counting the number of eigenvalues of the Hessian of the gravitational potential above a given threshold \citep[see also][]{forero2009}. In particular, the threshold adopted was chosen to obtain a volume filling fraction (VFF) of voids of about $70\%$ as done by \cite{2014MNRAS.445..988N} for their reconstruction on the local universe based on peculiar velocity fields. As a result, the corresponding VFFs of sheets, filaments and knots are uniquely determined by this choice.

 % 
% 
%@ARTICLE{2009MNRAS.396.1815F,
%   author = {{Forero-Romero}, J.~E. and {Hoffman}, Y. and {Gottl{\"o}ber}, S. and 
%{Klypin}, A. and {Yepes}, G.},
%    title = "{A dynamical classification of the cosmic web}",
%  journal = {\mnras},
%archivePrefix = "arXiv",
%   eprint = {0809.4135},
% keywords = {methods: numerical , cosmology: large-scale structure of Universe},
%     year = 2009,
%    month = jul,
%   volume = 396,
%    pages = {1815-1824},
%      doi = {10.1111/j.1365-2966.2009.14885.x},
%   adsurl = {http://adsabs.harvard.edu/abs/2009MNRAS.396.1815F},
%  adsnote = {Provided by the SAO/NASA Astrophysics Data System}
%}  
% 
%@ARTICLE{2014MNRAS.445..988N,
%   author = {{Nuza}, S.~E. and {Kitaura}, F.-S. and {He{\ss}}, S. and {Libeskind}, N.~I. and 
%{M{\"u}ller}, V.},
%    title = "{The cosmic web of the Local Universe: cosmic variance, matter content and its relation to galaxy morphology}",
%  journal = {\mnras},
%archivePrefix = "arXiv",
%   eprint = {1406.1004},
% keywords = {methods: numerical, methods: observational, galaxies: general, cosmology: theory, large-scale structure of Universe},
%     year = 2014,
%    month = nov,
%   volume = 445,
%    pages = {988-1001},
%      doi = {10.1093/mnras/stu1746},
%   adsurl = {http://adsabs.harvard.edu/abs/2014MNRAS.445..988N},
%  adsnote = {Provided by the SAO/NASA Astrophysics Data System}
%}
% 
% 
%@ARTICLE{2012MNRAS.425.2443K,
%   author = {{Kitaura}, F.-S. and {Angulo}, R.~E.},
%    title = "{Linearization with cosmological perturbation theory}",
%  journal = {\mnras},
%archivePrefix = "arXiv",
%   eprint = {1111.6617},
% keywords = {catalogues, galaxies: clusters: general, galaxies: statistics, large-scale structure of Universe, catalogues, galaxies: clusters: general, galaxies: statistics, large-scale structure of Universe},
%     year = 2012,
%    month = oct,
%   volume = 425,
%    pages = {2443-2454},
%      doi = {10.1111/j.1365-2966.2012.21614.x},
%   adsurl = {http://adsabs.harvard.edu/abs/2012MNRAS.425.2443K},
%  adsnote = {Provided by the SAO/NASA Astrophysics Data System}
%}
% 
% 
%@ARTICLE{2007MNRAS.375..489H,
%   author = {{Hahn}, O. and {Porciani}, C. and {Carollo}, C.~M. and {Dekel}, A.},
%    title = "{Properties of dark matter haloes in clusters, filaments, sheets and voids}",
%  journal = {\mnras},
%   eprint = {astro-ph/0610280},
% keywords = {methods: N-body simulations , galaxies: haloes , cosmology: theory , dark matter , large-scale structure of Universe},
%     year = 2007,
%    month = feb,
%   volume = 375,
%    pages = {489-499},
%      doi = {10.1111/j.1365-2966.2006.11318.x},
%   adsurl = {http://adsabs.harvard.edu/abs/2007MNRAS.375..489H},
%  adsnote = {Provided by the SAO/NASA Astrophysics Data System}
%}
% 
% 
%@ARTICLE{1974FizSz..24..304Z,
%   author = {{Zel'dovich}, Y.~B.},
%    title = "{The structure of the universe. A survey of contemporary cosmology.}",
%  journal = {Fizika Sz.},
%     year = 1974,
%   volume = 24,
%    pages = {304-308},
%   adsurl = {http://adsabs.harvard.edu/abs/1974FizSz..24..304Z},
%  adsnote = {Provided by the SAO/NASA Astrophysics Data System}
%